\chapter{Project Plan}

\section{Development Process}\label{Process}

In developing the BALL language, we followed the suggestion presented
by Prof. Aho during class: make a working prototype, and incrementally
add enhancements. During the early parts of the development, more
attention is focused on adding new features than testing, however each
team member must create his own test case with the changes he made.

Whenever a part of the language is finished, it is tested and reviewed
by the team. Feedback is then relayed to the team member responsible
for the subsystem (usually the one who merged the latest version to
trunk, as many modules are jointly developed by the team). There is
less of a focus on the traditional team roles, rather each team member
focuses on developing the compiler while also doing their job in team
administration.

\pagebreak
\section{Style Sheet}\label{Style}

\codelisting{StyleSheet.java}{StyleSheet.java}{../docs/StyleSheet.java}

\pagebreak

\section{Timeline \& Logs}\label{TimelineLogs}
\subsection{Timeline of BALL Development}

\begin{center}
\tablefirsthead{%
  \hline
  Date & Event \\
  \hline
}
\tablehead{%
  \hline
  \multicolumn{2}{|l|}{\footnotesize\sl continued from previous page}\\
  \hline
  Date & Event \\
  \hline
}
\tabletail{%
  \hline
  \multicolumn{2}{|r|}{\footnotesize\sl continued on next page}\\
  \hline}
\tablelasttail{\hline}
\bottomcaption{SVN commit log}
\small

\begin{supertabular}{|c|p{5in}|}
9/18/09, 11:31a & Team Llamamelon Born.\\
9/21/09, 5:15p & First Team Llamamelon Meeting; Team Llamamelon
named. Sam pioneers the language purpose: the simulation of baseball
games.\\
9/21/09, 8:02p & Team Llamamelon announced to PLT Teaching Staff.\\
10/2/09, 4:42p & Alfred V. Aho tentatively approves BALL programming
language.\\
10/5/09, 1:04a & Buzzwords for initial Whitepaper finalized.\\
10/7/07, 3:17p & Final Whitepaper completed.\\
10/23/09, 2:52p & Sample programs designed, BALL envisioned to be a
scripting language \`{a} la AWK.\\
10/27/09, 4:14p & Google Code Group Created. Jordan initializes
Subversion repository and trains team for first 'commit.'\\
10/27/09, 5:30p & First major Team Llamamelon Meeting; language
roadmap created.\\
10/29/09, 10:30a & First meeting with Sinan; BALL changes from
scripted language to compiled language, based on Java.\\
10/30/09, 1:27p & Nathan creates first example program; the new
language roadmap is born.\\
10/31/09, 10:42p & Cipta begins creating the grammar, laying the
groundwork for BALL.\\
11/3/09, 1:05p & Language Reference Manual and Tutorial begun.\\
11/3/09, 4:43p & Language Grammar completed, Daniel creates and
finalizes complex simulation to show off the powerful grammar.\\
11/4/09, 10:26a & Final Language Reference Manual and Tutorial
completed.\\
11/15/09, 3:00p & First semi-weekly meeting. Full grammar implemented
using JFlex/BYACC-J. Lexer created a DFA!\\
11/18/09, 8:00p & Second semi-weekly meeting. Play.ball designed,
intermediate code generation begins.\\
11/19/09, 11:00a & Second meeting with Sinan; Language Reference
Manual and Tutorial approved; BALL is in full motion!\\
11/27/09, 12:43a & Play.Ball compiles and runs: ``Play Ball!'' [Hello
  world equivalent.]\\
12/3/09, 3:45p & Team Loader created; teams can now load into the
PlayerObj and TeamObj constructs.\\
12/5/09, 11:46a & Major all-day meeting; BALL presentation created and
beautified.\\
12/7/09, 4:30p & BALL presented to PLT. There was much rejoicing.\\
12/10/09, 4:37a & SymbolTable embraced as major component of BALL.\\
12/10/09, 3:02a & Simfunc.ball compiles and runs: simfunctions and
activate work.\\
12/12/09, 10:00a & Major all-day meeting; Much of language completed
during this time.\\
12/12/09, 10:48a & TeamStat and PlayerStat methods added: stats work
throughout BALL.\\
12/12/09, 7:59p & Lists become typed, major change in the language.\\
12/12/09, 8:55p & Logical, Boolean, Comparison expressions completed.\\
12/15/09, 12:56a & \texttt{where}, \texttt{from}, and \texttt{any} keywords working.\\
12/16/09, 2:02a & \texttt{'s} keyword works for all attributes, stats, and
lists inherent in team and player objects.\\
12/16/09, 6:08p & Unary operators completed.\\
12/17/09, 6:27p & Realized the class number is W4115, not
W4119. [Seriously.]\\
12/20/09, 10:30a & Final Major all-day meeting: everything completed.\\
12/20/09, 8:04p & Everything in the language works. Final presentation
compilation begun.\\
Now & Final presentation completed.\\

\end{supertabular}

\normalsize
\end{center}

\pagebreak
\subsection{Project Log}

\input{projlog.tex}
